---
title: "What Does Music Look Like?"
format: html
execute:
  echo: false
  warning: false
  message: false
---

We normally experience music through sound. What happens when we turn the visual artwork surrounding music into data?

```{r}
#| label: setup
library(tidyverse)

# Load the verified albums dataset with visual features
albums <- read_csv("data/album_features.csv", show_col_types = FALSE)
```

## Visual Map: Brightness vs. Saturation

If we cannot hear the music, can we see patterns in its artwork? Here we plot each cover's **Brightness** (dark to light) against its **Saturation** (grayscale to vivid). Hover over any point to inspect the album.

```{r}
#| label: visual-map
#| fig-width: 9
#| fig-height: 5.5

library(plotly)

p <- ggplot(
  albums,
  aes(
    x = brightness,
    y = saturation,
    color = genre,
    text = paste0(
      "<b>", title, "</b><br>",
      artist, " (", year, ")<br>",
      "Genre: ", genre, "<br>",
      "Brightness: ", scales::percent(brightness, accuracy = 0.1), "<br>",
      "Saturation: ", scales::percent(saturation, accuracy = 0.1)
    )
  )
) +
  geom_point(size = 3.5, alpha = 0.85) +
  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(
    title = "Visual Map: Brightness vs. Saturation by Genre",
    x = "Brightness (0% = Dark, 100% = Light)",
    y = "Saturation (0% = Muted, 100% = Vivid)",
    color = "Genre"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

ggplotly(p, tooltip = "text")
```


## Album Explorer

Explore album covers across genres and decades.

```{r}
#| label: album-grid
#| results: asis

cards <- albums |>
  pmap_chr(\(title, artist, genre, year, decade, cover_url, ...) {
    paste0(
      '<div class="grid-card">',
      '<img src="', cover_url, '" alt="', htmltools::htmlEscape(title), '" loading="lazy">',
      '<div class="grid-info">',
      '<strong>', htmltools::htmlEscape(title), '</strong>',
      '<span>', htmltools::htmlEscape(artist), '</span>',
      '<small>', year, ' &bull; ', genre, '</small>',
      '</div></div>'
    )
  })

cat('<div class="album-grid">', cards, '</div>', sep = "\n")
```

```{=html}
<script>
  window.ALBUM_DATA = `r jsonlite::toJSON(albums)`;
</script>
<script src="timelapse.js"></script>
```
